\batcave: Bayesian Analysis Tools for Context-Aware Variant Evaluation

### 2019-09-10 NAR_DRAFT.tex
add r package name and remove comment

Here we present an algorithm for estimating the prior probability of mutation at a given site using the observed mutation spectrum of the tumor as well as its mutation rate, and show that the addition of this prior to the MuTect variant calling model produces a superior variant classifier in both simulated and real tumor data.
We provide a simple implementation in R that takes MuTect caller output as input, and returns the posterior probability that a site is variant for every site observed by MuTect. \bkmcomment{batcave,batcaver,batcaveR}

@@

\subsection{Real Data}
We analyzed two real data sets, one from an acute myeloid leukemia, and one from a multi-region sequencing experiment in breast cancer \cite{Griffith2015,Shi2018}.
We downloaded the normal and primary tumor bam files generated for the whole genome sequences in \citet{Griffith2015} from dbGaP accession number phs000159.v8.p4.
\citet{Griffith2015} generated a platinum set of variant calls for this primary tumor, which we obtained from supplementary materials, and used for our true positive dataset.
The whole exome sequencing bam files generated by \citep{Shi2018} were downloaded from NCBI SRA no. SRP070662.
\citet{Shi2018} generated a gold set of variant calls for each tumor region sequenced, which we obtained from supplementary materials and used for our true positive dataset.
We called variants using Mutect 1.1.7 with the same parameters as described above, except that both of these data sets were originally aligned to GRCH37, and we used that reference here.

>>>>

Here we present an algorithm for estimating the prior probability of mutation at a given site using the observed mutation spectrum of the tumor as well as its mutation rate, and show that the addition of this prior to the MuTect variant calling model produces a superior variant classifier in both simulated and real tumor data.
We provide a simple implementation in R, \texttt{batcaver} that takes MuTect caller output as input, and returns the posterior probability that a site is variant for every site observed by MuTect.

@@

\subsection{Real Data}
We analyzed two real data sets, one from an acute myeloid leukemia, and one from a multi-region sequencing experiment in breast cancer \cite{Griffith2015,Shi2018}.
We downloaded the normal and primary tumor bam files generated for the whole genome sequences in \citet{Griffith2015} from dbGaP accession number phs000159.v8.p4.
\citet{Griffith2015} generated a platinum set of variant calls for this primary tumor, which we obtained from supplementary materials, and used for our true positive dataset.
The whole exome sequencing bam files generated by \citet{Shi2018} were downloaded from NCBI SRA no. SRP070662.
\citet{Shi2018} generated a gold set of variant calls for each tumor region sequenced, which we obtained from supplementary materials and used for our true positive dataset.
We called variants using Mutect 1.1.7 with the same parameters as described above, except that both of these data sets were originally aligned to GRCH37, and we used that reference here.
